Different forms of uncertainty differently affect the evolution of social learning

### 2022-08-11 ms.tex
minor edits

Increased selection set size $B$ led nearly uniformly to more prevalent social
learning, with all other uncertainty parameters held constant. This is first 
because it is more difficuilt to find the optimal behavior with a larger $B$.
% more options present a more difficult problem for trial-and-error learning within
% a lifespan. 
This effect is amplified because agents in a world with larger $B$ are more
cognitively flexible compared to those in a smaller $B$ world. 
When $B$ is larger, agents are relatively more likely to choose an alternative to
their current known optimum. %is easier to correct for larger $B$ relative to smaller $B$.  
To understand this, consider the case where a social learner child learns
that behavior 1 has an expected payoff of 0.8, and the rest have an expected
payoff of 0, but the optimal behavior changed between generations. Assume
the child tries behavior 1, but it pays off 0.
If $B=2$, then at the next within-generation time step, $\ledger = (0.4, 0.0)$. The probability
the agent explores behavior 2 is now 0.02. However, if $B=10$ the agent has
$\ledger = (0.4, 0.0, \ldots, 0.0)$; the probability
the agent explores one of the nine alternative behaviors increases to 0.14.
If the agent again tries behavior 1, the probability of exploring alternatives
with $B=2$ is 0.21, but with $B=10$ the probability of exploring alternatives
greatly increases to 0.71. This explanation does not apply when $L=1$, where
individual learning does not occur. %social learning is the only possible form of learning.

@@

Our modeling choices were judiciously made to address the main question of how
uncertainty affects the evolution of social learning among populations of intelligent
autonomous problem solvers.  Our model necessarily made simplifying assumptions that,
while empirically justified across taxa, may not always hold in more
specific situations. First, note that we only
considered success-biased learning, although conformity is sometimes the dominant
social learning mechanism~\cite{Muthukrishna2016a,Smaldino2018b}.
Conformity may lead to more drift overall, since success-biased teacher selection, which
we assumed, maximizes the potential benefit to social learners. We assumed that the
number of behaviors and their payoffs were constant, but this fails to account for
evolutionary feedback which creates new behavioral opportunities as time progresses, e.g.\
via niche construction~\cite{Smaldino2012a,Heras-Escribano2020} or cumulative cultural
evolution supported by dynamic network structure~\cite{Smolla2019,Derex2020}. 
Group structure and processes such as homophily
and outgroup aversion could inhibit the evolution of social learning since 
successful outgroup teachers
could be cut off from ingroup learners~\cite{Jackson2012}. Finally, our individual
learning mechanism, softmax search, is underpowered compared to human individual learning; 
more powerful individual learning could further support social learning, and
this might be essential to accurate modeling of human behavior~\cite{Schulz2020a,Wu2022}.
These considerations reveal new opportunities to further develop a more nuanced 
theory of social learning by adjusting or extending the present model for other contexts.

>>>>

Increased selection set size $B$ led nearly uniformly to more prevalent social
learning, with all other uncertainty parameters held constant. This is first 
because it is more difficuilt to find the optimal behavior with a larger $B$.
% more options present a more difficult problem for trial-and-error learning within
% a lifespan. 
This effect is amplified because agents in a world with larger $B$ are more
cognitively flexible compared to those in a smaller $B$ world. 
When $B$ is larger, agents are relatively more likely to choose an alternative to
their current known optimum. %is easier to correct for larger $B$ relative to smaller $B$.  
To understand this, consider the case where a social learner child learns
that behavior 1 has an expected payoff of 0.8, and the rest have an expected
payoff of 0, but the optimal behavior changed between generations. Assume
the child tries behavior 1, but it pays off 0.
If $B=2$, then at the next within-generation time step, $\ledger = (0.4, 0.0)$. The probability
the agent explores behavior 2 is now 0.02. However, if $B=10$ the agent has
$\ledger = (0.4, 0.0, \ldots, 0.0)$; the probability
the agent explores one of the nine alternative behaviors increases to 0.14.
If the agent again tries behavior 1 and it pays 0, the probability of exploring alternatives
with $B=2$ is 0.21, but with $B=10$ the probability of exploring alternatives
greatly increases to 0.71. \mt{I double checked these calculations, but would appreciate if
someone triple checked it}. This explanation does not apply when $L=1$, where
individual learning does not occur.

@@

Our modeling choices were judiciously made to address the main question of how uncertainty
affects the evolution of social learning in general; however there exist alternative
empirically-valid choices that could help further develop a more nuanced theory of social
learning.  Our model necessarily made simplifying assumptions that may not always hold in more
specific situations, though they are empirically justified to be general across taxa. First,
note that we only considered success-biased learning, although conformity is sometimes the
dominant social learning mechanism~\cite{Muthukrishna2016a,Smaldino2018b}.  Conformity may lead
to more drift overall, since success-biased teacher selection, which we assumed, maximizes the
potential benefit to social learners. We assumed that the number of behaviors and their payoffs
were constant, but this fails to account for evolutionary feedback which creates new behavioral
opportunities as time progresses, e.g.\ via niche
construction~\cite{Smaldino2012a,Heras-Escribano2020} or cumulative cultural evolution
supported by dynamic network structure~\cite{Smolla2019,Derex2020}.  Group structure and
processes such as homophily and in-group bias/out-group aversion could inhibit the evolution of
social learning since successful outgroup teachers could be cut off from in-group
learners~\cite{Jackson2012}. We assumed agents choose behaviors via softmax search, but this is
an underpowered algorithm compared to human individual learning~\cite{Schulz2020a,Wu2022}.
More powerful individual learning could further support social learning.  These considerations
reveal new opportunities to further by adjusting or extending the present model for other
contexts.
